---
title: "Bootstrapped estimation of D scores for individual participants"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

# Overview 

IRAP trial-type *D* scores are calculated from an average of only 18 pairs of reaction times. This would be deemed as far too low anywhere else in the literature on reaction time based tasks. The implications of this can be seen in how poorly estimated any one IRAP D score is. We can observe this by bootstrapping reaction times for each participant's *D* scores and noting how wide their confidence intervals are.

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

```{r}

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
library(boot)
library(parallel)
library(bayestestR)
library(patchwork)
library(mdthemes)
library(lme4)
library(sjPlot)
library(emmeans)
library(ggstance)
library(janitor)
# library(merTools) called via merTools:: to avoid namespace collisions between MASS and dplyr


# set seed for reproducibility
set.seed(42)

# number of bootstraps
n_boots <- 5000 

# options
options(knitr.table.format = "html") # necessary configuration of tables

# disable scientific notation
options(scipen = 999) 


# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  require(janitor)
  df %>% mutate_if(is.numeric, janitor::round_half_up, digits = n_digits)
}


# create necessary directories
dir.create("../data/processed")
dir.create("../data/results")
dir.create("models")


# get data from evaluative IRAPs
data_trial_level <- read_csv("../data/raw/data_trial_level.csv") %>%
  filter(timepoint == "baseline" & (age >= 18 | is.na(age)))

# outliers
data_outliers <- data_trial_level %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  select(unique_id, domain, mean_rt) %>%
  mutate(median_mean_rt = median(mean_rt, na.rm = TRUE),
         mad_mean_rt = mad(mean_rt, na.rm = TRUE)) %>%
  # exclude median +- 2MAD
  mutate(rt_outlier = ifelse(mean_rt < median_mean_rt-mad_mean_rt*2 |
                               mean_rt > median_mean_rt+mad_mean_rt*2, TRUE, FALSE)) %>%
  filter(rt_outlier == FALSE) %>%
  select(unique_id, rt_outlier) %>%
  full_join(data_trial_level, by = "unique_id") %>%
  mutate(rt_outlier = ifelse(is.na(rt_outlier), TRUE, rt_outlier))

data_outliers_removed <- data_outliers %>%
  filter(rt_outlier == FALSE)

# trim RTs>10000 ms, as part of D scoring
data_trimmed <- data_outliers_removed %>%
  select(unique_id, domain, trial_type, rt, block_type) %>%
  filter(rt <= 10000)

```

# Sample descriptives

```{r}

data_outliers %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  count(rt_outlier) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

data_descriptives <- data_outliers_removed %>%
  distinct(unique_id, .keep_all = TRUE)

data_descriptives %>%
  count(domain) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

data_descriptives %>%
  count(domain) %>%
  summarize(total_n = sum(n),
            min_n_per_domain = min(n),
            max_n_per_domain = max(n),
            mean_n_per_domain = round(mean(n, na.rm = TRUE), 2),
            sd_n_per_domain = round(sd(n, na.rm = TRUE), 2)) %>%
  gather() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

data_descriptives %>%
  summarize(min_age  = round(min(age, na.rm = TRUE), 2),
            max_age  = round(max(age, na.rm = TRUE), 2),
            mean_age = round(mean(age, na.rm = TRUE), 2),
            sd_age   = round(sd(age, na.rm = TRUE), 2)) %>%
  gather() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

data_descriptives %>%
  count(gender) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Distribution of RTs

In order to convey how small these effects are and what they're estimated from for a given participant.

## All RTs

```{r}

data_trimmed %>%
  mutate(trial_type = case_when(trial_type == "tt1" ~ "Trial type 1",
                                trial_type == "tt2" ~ "Trial type 2",
                                trial_type == "tt3" ~ "Trial type 3",
                                trial_type == "tt4" ~ "Trial type 4")) %>%
  ggplot(aes(rt, fill = block_type)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ trial_type, ncol = 1) +
  mdthemes::md_theme_linedraw() +
  scale_fill_viridis_d(begin = 0.3, end = 0.7) +
  ylab("Frequency") +
  xlab("Reaction time (ms)") +
  labs(fill = "Block type")

# data_trimmed %>%
#   group_by(unique_id, trial_type) %>%
#   summarize(mean_rt_con = mean(rt[block_type == "con"], na.rm = TRUE),
#             mean_rt_incon = mean(rt[block_type == "incon"], na.rm = TRUE),
#             sd_rt_con = sd(rt[block_type == "con"], na.rm = TRUE),
#             sd_rt_incon = sd(rt[block_type == "con"], na.rm = TRUE),
#             sd_rt = sd(rt, na.rm = TRUE),
#             .groups = "drop") %>%
#   mutate(diff_mean_rt = mean_rt_incon - mean_rt_con) %>%
#   select(-unique_id) %>%
#   group_by(trial_type) %>%
#   summarize_all(median) %>%
#   round_df(0) %>%
#   select(trial_type, 
#          median_mean_rt_con = mean_rt_con, 
#          median_mean_rt_incon = mean_rt_incon, 
#          median_diff_mean_rt = diff_mean_rt, 
#          median_sd_rt_con = sd_rt_con, 
#          median_sd_rt_incon = sd_rt_incon, 
#          median_sd_rt = sd_rt) %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## Single participant with median D score (D = 0.10)

```{r}

data_eg_1 <- data_trimmed %>%
  filter(unique_id == "ian death emma_43" & trial_type == "tt1") 
  #filter(unique_id == "ian death emma_15" & trial_type == "tt1") 
  #filter(unique_id == "ian death june_57" & trial_type == "tt1") 
  

#dat <- data_trimmed %>%
data_eg_1 %>%  
  group_by(unique_id, trial_type) %>%
  summarize(mean_rt_con = mean(rt[block_type == "con"], na.rm = TRUE),
            mean_rt_incon = mean(rt[block_type == "incon"], na.rm = TRUE),
            sd_rt = sd(rt, na.rm = TRUE)) %>%
  mutate(D = (mean_rt_incon - mean_rt_con)/sd_rt)

ggplot(data_eg_1, aes(rt, block_type)) +
  geom_point(alpha = 0.5, position = position_jitter(height = 0.2, 
                                                     width = 0,
                                                     seed = 43)) +
  mdthemes::md_theme_linedraw() +
  labs(x = "Reaction time (ms)",
       y = "Block type") 

```

# Calculate scores

## Bootstrap 95% CIs on *D* scores

Calculated by domain and trial-type. 1+ hour runtime on a good PC.

```{r}

# bootstrapping has a long execution time, so load saved values if they've already been calculated
if(file.exists("../data/processed/data_estimates_D.csv")) {
  
  data_estimates_D <- read_csv("../data/processed/data_estimates_D.csv")
  
} else {
  
  D_score <- function(data, i) {
    data_with_indexes <- data[i,] # boot function requires data and index
    mean_con <- mean(data_with_indexes$rt[data_with_indexes$block_type == "con"], na.rm = TRUE)
    mean_incon <- mean(data_with_indexes$rt[data_with_indexes$block_type == "incon"], na.rm = TRUE)
    sd <- sd(data_with_indexes$rt, na.rm = TRUE)
    D <- (mean_incon - mean_con) / sd
    return(D)
  }
  
  bootstrap_D_score <- function(data){
    
    require(dplyr)
    require(boot)
    
    fit <- 
      boot::boot(data      = data, 
                 statistic = D_score, 
                 R         = n_boots, 
                 sim       = "ordinary", 
                 stype     = "i",
                 parallel  = "multicore", 
                 ncpus     = parallel::detectCores())
    
    results <- boot::boot.ci(fit, conf = 0.95, type = c("norm", "basic", "perc", "bca"))
    
    output <- 
      tibble(method   = c("normal", "basic", "percent", "bca"),
             estimate = rep(fit$t0, 4),
             ci_lower = c(results$normal[2], results$basic[4], results$percent[4], results$bca[4]),
             ci_upper = c(results$normal[3], results$basic[5], results$percent[5], results$bca[5]))
    
    return(output)
  }
  
  # bootstrap D scores 
  data_estimates_D <- data_trimmed %>%
    select(unique_id, domain, trial_type, rt, block_type) %>%
    group_by(unique_id, domain, trial_type) %>%
    do(bootstrap_D_score(data = .)) %>%
    ungroup() %>%
    mutate(sig = ifelse((ci_lower < 0 & ci_upper < 0) | (ci_lower > 0 & ci_upper > 0), TRUE, FALSE),
           ci_width = ci_upper - ci_lower) %>%
    round_df(3)
  
  # save to disk
  write_csv(data_estimates_D, "../data/processed/data_estimates_D.csv")

}

```

## Bootstrapped 95% CIs on PI scores 

Calculated by domain and trial-type. 1+ hour runtime on a good PC.

```{r}

# bootstrapping has a long execution time, so load saved values if they've already been calculated
if(file.exists("../data/processed/data_estimates_PI.csv")) {
  
  data_estimates_PI <- read_csv("../data/processed/data_estimates_PI.csv")
  
} else {
  
  # Fast calculation of the A statistic - code from Ruscio (2008) supplementary materials
  PI_score <- function(data, i) {
    data_with_indexes <- data[i,] # boot function requires data and index
    x  <- na.omit(data_with_indexes$rt[data_with_indexes$block_type == "incon"])
    y  <- na.omit(data_with_indexes$rt[data_with_indexes$block_type == "con"])
    nx <- length(x)
    ny <- length(y)
    rx <- sum(rank(c(x, y))[1:nx])
    PI <- (rx / nx - (nx + 1) / 2) / ny
    return(PI)
  }
  
  bootstrap_PI_score <- function(data){
    
    require(dplyr)
    require(boot)
    
    fit <- 
      boot::boot(data      = data, 
                 statistic = PI_score, 
                 R         = n_boots, 
                 sim       = "ordinary", 
                 stype     = "i",
                 parallel  = "multicore", 
                 ncpus     = parallel::detectCores())
    
    results <- boot::boot.ci(fit, conf = 0.95, type = c("norm","basic", "perc", "bca"))
    
    output <- 
      tibble(method   = c("normal", "basic", "percent", "bca"),
             estimate = rep(fit$t0, 4),
             ci_lower = c(results$normal[2], results$basic[4], results$percent[4], results$bca[4]),
             ci_upper = c(results$normal[3], results$basic[5], results$percent[5], results$bca[5]))
    
    return(output)
  }
  
  # bootstrap PI scores 
  data_estimates_PI <- data_outliers_removed %>%
    group_by(unique_id, domain, trial_type) %>%
    do(bootstrap_PI_score(data = .)) %>%
    ungroup() %>%
    mutate(sig = ifelse((ci_lower < 0.50 & ci_upper < 0.50) | (ci_lower > 0.50 & ci_upper > 0.50), TRUE, FALSE),
           ci_width = ci_upper - ci_lower) %>%
    round_df(3)
  
  # save to disk
  write_csv(data_estimates_PI, "../data/processed/data_estimates_PI.csv")
  
}

```

## Plots

By domain, using normal bootstrapping

```{r fig.height=8, fig.width=8}

p_cis_by_domain <- 
  data_estimates_D %>%
  filter(method == "basic") %>%
  mutate(domain = str_replace(domain, "Personality - ", "Big5: "),
         domain = str_replace(domain, "Stigma - ", "Stigma: ")) %>%
  arrange(estimate) %>%
  group_by(domain) %>%
  mutate(ordered_id = row_number()/n()) %>%
  ungroup() %>%
  ggplot() +
  geom_linerange(aes(x = ordered_id, ymin = ci_lower, ymax = ci_upper, color = sig),
                 alpha = 1) +
  geom_point(aes(ordered_id, estimate), size = 0.5, shape = "square") +
  geom_hline(yintercept = 0, linetype = "dotted") +
  mdthemes::md_theme_linedraw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "top") +
  scale_color_viridis_d(end = 0.6, direction = -1) +
  xlab("Ranked participant") +
  ylab("*D* score") +
  labs(color = "95% CI excludes zero point") + 
  facet_wrap(~ domain, ncol = 5)

p_cis_by_domain

```

# CI widths

Widths cant be directly compared between *D* and PI as they have different ranges, so *D* scores only.

Not meta analyzed as extreme skew in data means that residuals are very non-normal, violating assumptions and underestimating MAP estimates. Instead I simply present MAP estimates. 

## MAP-MAP width

Most probable estimate among the most probable estimates

```{r}

data_map_ci_widths <- data_estimates_D %>%
  mutate(method = fct_relevel(method, "bca", "basic", "normal", "percent")) %>%
  group_by(domain, trial_type, method) %>%
  do(point_estimate(.$ci_width, centrality = "MAP")) %>%
  ungroup()

data_map_ci_widths %>%
  group_by(method) %>%
  summarize(map_map = point_estimate(MAP, centrality = "MAP"),
            min_map = min(MAP),
            max_map = max(MAP)) %>%
  unnest(map_map) %>%
  rename(MAP_MAP = MAP) %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## MAP CI width

By domain and trial type using basic bootstrapping, on the basis that it has the best performance for % of non-zero *D* scores (further below).

```{r fig.height=16, fig.width=16}

data_map_ci_widths %>%
  filter(method == "basic") %>%
  pivot_wider(names_from = trial_type, values_from = MAP) %>%
  round_df(2) %>%
  dplyr::select(-method) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## Plot by domain and trial type

```{r fig.height=10, fig.width=8}

data_ci_width_map_D <- data_estimates_D %>%
  group_by(domain, trial_type, method) %>%
  #summarize(median = median(ci_width), .groups = "drop") %>%
  do(point_estimate(.$ci_width, centrality = "MAP")) %>%
  ungroup() %>%
  mutate(MAP = round_half_up(MAP, 3),
         method = fct_relevel(method, "bca", "basic", "normal", "percent"),
         method = fct_rev(method),
         trial_type = case_when(trial_type == "tt1" ~ "Trial type 1",
                                trial_type == "tt2" ~ "Trial type 2",
                                trial_type == "tt3" ~ "Trial type 3",
                                trial_type == "tt4" ~ "Trial type 4"),
         trial_type = fct_relevel(trial_type, "Trial type 1", "Trial type 2", "Trial type 3", "Trial type 4")) %>%
  #mutate(domain = fct_reorder(domain, MAP, .fun = min)) %>%
  #arrange(domain) %>%
  mutate(domain = fct_rev(domain))

# save to disk
write_csv(data_ci_width_map_D, "../data/results/data_ci_width_map_D.csv")

# plot
p_ci_widths <- 
  ggplot(data_ci_width_map_D, aes(MAP, domain, color = method, shape = method)) + 
  geom_point(position = position_dodge(width = 0.8)) +
  scale_shape_manual(labels = c("Percentile", "Normal", "Basic", "BCA"), values = c(17, 16, 15, 19)) +
  scale_color_viridis_d(begin = 0.2, end = 0.8, labels = c("Percentile", "Normal", "Basic", "BCA"), direction = -1) +
  mdthemes::md_theme_linedraw() +
  facet_wrap(~ trial_type, ncol = 4, nrow = 1) +
  labs(x = "Highest probability (MAP) 95% CI width",
       y = "",
       color = "Bootstrapping method",
       shape = "Bootstrapping method") + 
  theme(legend.position = "top")
  
p_ci_widths

```

# Proportion different from zero

## Calculate scores

```{r}

data_diff_zero <- 
  bind_rows(mutate(data_estimates_D, DV_type = "*D* scores"),
            mutate(data_estimates_PI, DV_type = "PI scores")) %>%
  mutate(domain = as.factor(domain),
         method = fct_relevel(method, "bca", "basic", "normal", "percent"),
         method = fct_rev(method),
         trial_type = case_when(trial_type == "tt1" ~ "Trial type 1",
                                trial_type == "tt2" ~ "Trial type 2",
                                trial_type == "tt3" ~ "Trial type 3",
                                trial_type == "tt4" ~ "Trial type 4"),
         trial_type = fct_relevel(trial_type, "Trial type 1", "Trial type 2", "Trial type 3", "Trial type 4")) %>%
  group_by(domain, trial_type, method, DV_type) %>%
  summarize(proportion_diff_zero = mean(sig),
            variance = plotrix::std.error(sig)^2,
            .groups = "drop") %>%
  mutate(variance = ifelse(variance == 0, 0.000001, variance))  # model cannot be run on zero variance, so offset by a minuscule amount

# save to disk
write_csv(data_diff_zero, "../data/results/data_diff_zero.csv")

```

## Compare CI bootstrapping methods 

### Plot

```{r fig.height=10, fig.width=8}

p_diff_zero_method <- 
  data_diff_zero %>%
  filter(DV_type == "*D* scores") %>%
  mutate(domain = fct_reorder(domain, proportion_diff_zero, .fun = mean)) %>%
  ggplot(aes(proportion_diff_zero, domain, color = method, shape = method)) +
  geom_linerangeh(aes(xmin = proportion_diff_zero - sqrt(variance)*1.96,
                      xmax = proportion_diff_zero + sqrt(variance)*1.96),
                  position = position_dodge(width = 0.8)) + 
  geom_point(position = position_dodge(width = 0.8)) +
  scale_shape_manual(labels = c("Percentile", "Normal", "Basic", "BCA"), values = c(17, 16, 15, 19)) +
  scale_color_viridis_d(begin = 0.2, end = 0.8, labels = c("Percentile", "Normal", "Basic", "BCA"), direction = -1) +
  mdthemes::md_theme_linedraw() +
  facet_wrap(~ trial_type, ncol = 4) +
  labs(x = "Proportion of scores different from zero point",
       y = "",
       color = "Bootstrapping method",
       shape = "Bootstrapping method") + 
  theme(legend.position = "top")

p_diff_zero_method

```

### Model

```{r}

# fit model
fit_diff_zero_method <- 
  data_diff_zero %>%
  filter(DV_type == "*D* scores") %>%
  lmer(proportion_diff_zero ~ 1 + method + (1 | domain/trial_type),
       weights = 1/variance, 
       data = .)

# results table
tab_model(fit_diff_zero_method, 
          ci.hyphen = ", ")

# extract re Tau
results_re_tau_diff_zero_method <- fit_diff_zero_method %>%
  merTools::REsdExtract() %>%
  as_tibble(rownames = "trial_type") %>%
  rename(tau = value) 

# extract marginal means
results_emm_diff_zero_method <- 
  summary(emmeans(fit_diff_zero_method, ~ method)) %>%
  dplyr::select(method, estimate = emmean, se = SE, ci_lower = lower.CL, ci_upper = upper.CL) 

# combine
results_diff_zero_method <- results_emm_diff_zero_method %>%
  mutate(pi_lower = estimate - (1.96 * sqrt(se^2 + sum(results_re_tau_diff_zero_method$tau)^2)),  # as in metafor package's implementation of credibility intervals, see metafor::predict.rma.R 
         pi_upper = estimate + (1.96 * sqrt(se^2 + sum(results_re_tau_diff_zero_method$tau)^2)))

# plot
p_prop_nonzero_method <- 
  ggplot(results_diff_zero_method, aes(method, estimate, color = method, shape = method, group = method)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pi_upper), position = position_dodge(width = 0.25), linetype = "dotted") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size = 1, position = position_dodge(width = 0.25)) +
  geom_point(position = position_dodge(width = 0.25), size = 2.5) +
  mdthemes::md_theme_linedraw() +
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), labels = c("0.00<br/>(Worse)", "0.25", "0.50", "0.75", "1.00<br/>(Better)")) +
  scale_shape_manual(labels = c("Percentile", "Normal", "Basic", "BCA"), values = c(17, 16, 15, 19)) +
  scale_color_viridis_d(alpha = 1, begin = 0.2, end = 0.8, labels = c("Percentile", "Normal", "Basic", "BCA"), direction = -1) +
  labs(shape = "Bootstrapping method",
       color = "Bootstrapping method",
       x = "",
       y = "Proportion of participants with non-zero scores<br/>") + 
  theme(legend.position = "none") +
  coord_flip(ylim = c(0, 1))

p_prop_nonzero_method

results_diff_zero_method %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## Compare *D* vs PI scores

### Plot

```{r fig.height=7, fig.width=8}

# data_diff_zero %>%
#   filter(method %in% c("basic")) %>%
#   mutate(DV_type = case_when(DV_type == "*D* scores" ~ "D",
#                              DV_type == "PI scores" ~ "PI")) %>%
#   select(-variance) %>%
#   pivot_wider(names_from = c(DV_type, method),
#               values_from = proportion_diff_zero) %>%
#   round_df(2) %>%
#   arrange(domain, trial_type) %>%
#   janitor::clean_names() %>%
#   kable() %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

p_diff_zero_scoring <- 
  data_diff_zero %>%
  filter(method %in% c("basic")) %>%
  mutate(domain = fct_rev(factor(domain))) %>%
  ggplot(aes(proportion_diff_zero, domain, color = DV_type, shape = DV_type)) +
  geom_linerangeh(aes(xmin = proportion_diff_zero - sqrt(variance)*1.96,
                      xmax = proportion_diff_zero + sqrt(variance)*1.96),
                  position = position_dodge(width = 0.75)) + 
  geom_point(position = position_dodge(width = 0.75)) +
  scale_shape_manual(labels = c("*D* scores", "PI scores"), values = c(15, 16)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7, labels = c("*D* scores", "PI scores")) +
  mdthemes::md_theme_linedraw() +
  facet_wrap(~ trial_type, ncol = 4) +
  labs(x = "Proportion of scores different from zero point",
       y = "",
       color = "Scoring method",
       shape = "Scoring method") + 
  theme(legend.position = "top")

p_diff_zero_scoring

```

### Model

```{r}

# fit model
fit_diff_zero_scoring <- 
  data_diff_zero %>%
  filter(method %in% c("basic")) %>%
  lmer(proportion_diff_zero ~ 1 + DV_type + (1 | domain/trial_type),
       weights = 1/variance, 
       data = .)

# results table
tab_model(fit_diff_zero_scoring, 
          ci.hyphen = ", ")

# extract re Tau
results_re_tau_diff_zero_scoring <- fit_diff_zero_scoring %>%
  merTools::REsdExtract() %>%
  as_tibble(rownames = "trial_type") %>%
  rename(tau = value) 

# extract marginal means
results_emm_diff_zero_scoring <- 
  summary(emmeans(fit_diff_zero_scoring, ~ DV_type)) %>%
  dplyr::select(DV_type, estimate = emmean, se = SE, ci_lower = lower.CL, ci_upper = upper.CL)

# combine
results_diff_zero_scoring <- results_emm_diff_zero_scoring %>%
  mutate(pi_lower = estimate - (1.96 * sqrt(se^2 + sum(results_re_tau_diff_zero_scoring$tau)^2)),  # as in metafor package's implementation of prediction intervals, see metafor::predict.rma.R 
         pi_upper = estimate + (1.96 * sqrt(se^2 + sum(results_re_tau_diff_zero_scoring$tau)^2)))

# plot
p_prop_nonzero_scoring <- 
  ggplot(results_diff_zero_scoring, aes(DV_type, estimate, color = DV_type, shape = DV_type, group = DV_type)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pi_upper), position = position_dodge(width = 0.8), linetype = "dotted") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size = 1, position = position_dodge(width = 0.8)) +
  geom_point(position = position_dodge(width = 0.8), size = 2.5) +
  mdthemes::md_theme_linedraw() +
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), labels = c("0.00<br/>(Worse)", "0.25", "0.50", "0.75", "1.00<br/>(Better)")) +
  scale_color_viridis_d(alpha = 1, begin = 0.3, end = 0.7, labels = c("*D* scores", "PI scores")) +
  labs(shape = "Scoring method",
       color = "Scoring method",
       x = "",
       y = "Proportion of participants with non-zero scores<br/>") + 
  theme(legend.position = "none") +
  coord_flip(ylim = c(0, 1))

p_prop_nonzero_scoring

results_diff_zero_scoring %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Proportion different from one another

Within domain and trial type.

Note: Discriminability between a score and zero can be determined using the CI, because zero is a known value and only the score is measured with uncertainty. However, discriminability between two scores must take into account the uncertainty in the estimation of both scores. Weir (2005) argues that such an interval can be estimated by expanding the CIs by sqrt(2). Here I refer to these intervals as Discriminability Intervals (DIs).

## Calculate discriminability

Many have argued that the zero point is arbitrary and not a useful reference point. Instead of asking "what proportion of *D*/PI scores are different from zero?", we could also ask "what proportion of *D*/PI scores are different from one another?"

```{r}

# helper function to apply workflow to each resample
discriminability <- function(data, i) {
  
  data_with_indexes <- data[i,] # boot function requires data and index
  
  estimate <- data_with_indexes$estimate
  di_lower <- data_with_indexes$di_lower 
  di_upper <- data_with_indexes$di_upper

  n_estimate <- length(estimate)
  n_di_lower <- length(di_lower)
  n_di_upper <- length(di_upper)
  
  r_estimate <- sum(rank(c(estimate, di_lower))[1:n_estimate])
  r_di_upper <- sum(rank(c(di_upper, estimate))[1:n_di_upper])
  
  prob_estimate_inferior_to_di_lower <- 1 - (r_estimate / n_estimate - (n_estimate + 1) / 2) / n_di_lower
  prob_estimate_superior_to_di_upper <- 1 - (r_di_upper / n_di_upper - (n_di_upper + 1) / 2) / n_estimate
  
  probability_estimates_outside_cis <- (prob_estimate_inferior_to_di_lower + prob_estimate_superior_to_di_upper)
  
  return(probability_estimates_outside_cis)
  
}

bootstrap_discriminability <- function(data){
  
  require(dplyr)
  require(boot)
  
  fit <- 
    boot::boot(data      = data, 
               statistic = discriminability, 
               R         = n_boots,
               sim       = "ordinary", 
               stype     = "i",
               parallel  = "multicore", 
               ncpus     = parallel::detectCores())
  
  #results <- boot::boot.ci(fit, conf = 0.95, type = c("norm","basic", "perc", "bca"))
  results <- boot::boot.ci(fit, conf = 0.95, type = c("basic"))
  
  output <-
    tibble(
      method   = c("normal", "basic", "percent", "bca"),
      estimate = rep(fit$t0, 4),
      ci_lower = c(results$normal[2], results$basic[4], results$percent[4], results$bca[4]),
      ci_upper = c(results$normal[3], results$basic[5], results$percent[5], results$bca[5])
    )
  
  # output <- 
  #   tibble(
  #     method   = "bca",
  #     estimate = fit$t0,
  #     ci_lower = results$basic[4],
  #     ci_upper = results$basic[5]
  #   )
  
  return(output)
}

```

### *D* scores

```{r}

# bootstrapping has a long execution time, so load saved values if they've already been calculated
if(file.exists("../data/results/data_discriminability_D.csv")) {
  
  data_discriminability_D <- read_csv("../data/results/data_discriminability_D.csv")
  
} else {
  
  # bootstrap D scores 
  data_discriminability_D <- data_estimates_D |>
    # expand CIs by sqrt(2) to form discriminability intervals
    mutate(di_lower = estimate - (estimate - ci_lower)*sqrt(2),
           di_upper = estimate + (ci_upper - estimate)*sqrt(2)) |>
    select(unique_id, domain, trial_type, method, estimate, di_upper, di_lower) |>
    group_by(domain, trial_type, method) %>%
    do(bootstrap_discriminability(data = .)) |>
    ungroup() |>
    rename(proportion_discriminable = estimate) |>
    mutate(variance = ((ci_upper - ci_lower)/(1.96*2))^2,
           domain = as.factor(domain),
           method = fct_relevel(method, "bca", "basic", "normal", "percent"),
           trial_type = fct_relevel(trial_type, "tt1", "tt2", "tt3", "tt4"),
           DV_type = "*D* scores") |>
    round_df(4)

  # save to disk
  write_csv(data_discriminability_D, "../data/results/data_discriminability_D.csv")
  
}

```

### PI scores

```{r}

# bootstrapping has a long execution time, so load saved values if they've already been calculated
if(file.exists("../data/results/data_discriminability_PI.csv")) {
  
  data_discriminability_PI <- read_csv("../data/results/data_discriminability_PI.csv")
  
} else {
  
  # bootstrap D scores 
  data_discriminability_PI <- data_estimates_PI |>
    # expand CIs by sqrt(2) to form discriminability intervals
    mutate(di_lower = estimate - (estimate - ci_lower)*sqrt(2),
           di_upper = estimate + (ci_upper - estimate)*sqrt(2)) |>
    select(unique_id, domain, trial_type, method, estimate, di_upper, di_lower) |>
    group_by(domain, trial_type, method) %>%
    do(bootstrap_discriminability(data = .)) |>
    ungroup() |>
    rename(proportion_discriminable = estimate) |>
    mutate(variance = ((ci_upper - ci_lower)/(1.96*2))^2,
           domain = as.factor(domain),
           method = fct_relevel(method, "bca", "basic", "normal", "percent"),
           trial_type = fct_relevel(trial_type, "tt1", "tt2", "tt3", "tt4"),
           DV_type = "PI scores") |>
    round_df(4)

  # save to disk
  write_csv(data_discriminability_PI, "../data/results/data_discriminability_PI.csv")
  
}

```

## Compare CI bootstrapping methods

```{r}

# fit meta analytic model
fit_disciminability_method <- data_discriminability_D %>%
  lmer(proportion_discriminable ~ 1 + method + (1 | domain/trial_type), 
       #weights = 1/variance, 
       data = .)

data_discriminability_D %>%
  group_by(method) %>%
  summarize(mean_proportion_discriminable = mean(proportion_discriminable))

# results table
tab_model(fit_disciminability_method, 
          ci.hyphen = ", ")

# # extract re Tau
# results_re_tau_disciminability_method <- fit_disciminability_method %>%
#   merTools::REsdExtract() %>%
#   #as_tibble(rownames = "trial_type") %>%
#   as_tibble(rownames = "method") %>%
#   rename(tau = value) 
# 
# # extract marginal means
# results_emm_disciminability_method <- 
#   summary(emmeans(fit_disciminability_method, ~ 1)) %>%  # throws "system is computationally singular" error. ie estimates are identical. therefore taken from table
#   dplyr::select(method, estimate = emmean, se = SE, ci_lower = lower.CL, ci_upper = upper.CL)

```

- singular fit error; estimates are identical between methods, therefore no plot.

## Compare *D* vs PI scores

Model & plot

```{r}

# combine
data_discriminability_combined_scoring <- 
  bind_rows(data_discriminability_D,
            data_discriminability_PI) %>%
  filter(method == "basic") %>%
  mutate(trial_type = case_when(trial_type == "tt1" ~ "Trial type 1",
                                trial_type == "tt2" ~ "Trial type 2",	
                                trial_type == "tt3" ~ "Trial type 3",	
                                trial_type == "tt4" ~ "Trial type 4"),
         variance = ifelse(variance == 0, 0.0001, variance))

# fit meta analytic model
fit_disciminability_scoring <- 
  lmer(proportion_discriminable ~ 1 + DV_type + (1 | domain/trial_type), 
       weights = 1/variance, 
       data = data_discriminability_combined_scoring)

# results table
tab_model(fit_disciminability_scoring, 
          ci.hyphen = ", ")

# # plot RE effects
# plot_model(fit_disciminability_scoring,
#            colors = "bw",
#            type = "re") +
#   mdthemes::md_theme_linedraw() +
#   ggtitle("")

# extract re Tau
results_re_tau_disciminability_scoring <- fit_disciminability_scoring %>%
  merTools::REsdExtract() %>%
  as_tibble(rownames = "trial_type") %>%
  rename(tau = value) 

# extract marginal means
results_emm_disciminability_scoring <-
  summary(emmeans(fit_disciminability_scoring, ~ DV_type)) %>%
  dplyr::select(DV_type, estimate = emmean, se = SE, ci_lower = lower.CL, ci_upper = upper.CL) 

# combine
results_disciminability_scoring <- results_emm_disciminability_scoring %>%
  mutate(pi_lower = estimate - (1.96 * sqrt(se^2 + sum(results_re_tau_disciminability_scoring$tau)^2)),  # as in metafor package's implementation of credibility intervals, see metafor::predict.rma.R
         pi_upper = estimate + (1.96 * sqrt(se^2 + sum(results_re_tau_disciminability_scoring$tau)^2)))

# plot
p_prop_discriminable_scoring <-
  ggplot(results_disciminability_scoring, aes(DV_type, estimate, color = DV_type, shape = DV_type, group = DV_type)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pi_upper), position = position_dodge(width = 0.8), linetype = "dotted") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size = 1, position = position_dodge(width = 0.8)) +
  geom_point(position = position_dodge(width = 0.8), size = 2.5) +
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), labels = c("0.00<br/>(Worse)", "0.25", "0.50", "0.75", "1.00<br/>(Better)")) +
  scale_shape_manual(labels = c("*D* scores", "PI scores"), values = c(15, 16)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7, labels = c("*D* scores", "PI scores")) +
  mdthemes::md_theme_linedraw() +
  labs(x = "",
       y = "Proportion of participants who whose scores<br/>differ from one another<br/>",
       color = "Scoring method",
       shape = "Scoring method") +
  theme(legend.position = "none") +
  coord_flip(ylim = c(0, 1))

p_prop_discriminable_scoring 

results_disciminability_scoring %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# CI widths as a proportion of observed range

## Observed range

```{r}

data_estimates_D %>%
  mutate(method = fct_relevel(method, "bca", "basic", "normal", "percent")) %>%
  summarize(min_observed_Dscore = min(estimate),
            max_observed_Dscore = max(estimate)) %>%
  mutate(observed_range_Dscore = max_observed_Dscore - min_observed_Dscore) %>% 
  round_df(2) %>%
  gather() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## Calculate scores

```{r}

## calculate observed ranges 
observed_range_estimates_D <- data_estimates_D %>%
  group_by(domain, method, trial_type) %>%
  dplyr::summarize(min = min(estimate, na.rm = TRUE),
                   max = max(estimate, na.rm = TRUE),
                   .groups = "drop") %>%
  mutate(range = max - min) 

observed_range_estimates_PI <- data_estimates_PI %>%
  group_by(domain, method, trial_type) %>%
  dplyr::summarize(min = min(estimate, na.rm = TRUE),
                   max = max(estimate, na.rm = TRUE),
                   .groups = "drop") %>%
  mutate(range = max - min) 


# calculate CI / range 
data_ci_width_proportions_D <- data_estimates_D %>%
  # join this data into the original data
  full_join(observed_range_estimates_D, by = c("domain", "method", "trial_type")) %>%
  # calculate ci width as a proportion of observed range
  mutate(ci_width_proportion = ci_width / range) %>%
  mutate(DV_type = "*D* scores") 

data_ci_width_proportions_PI <- data_estimates_PI %>%
  # join this data into the original data
  full_join(observed_range_estimates_PI, by = c("domain", "method", "trial_type")) %>%
  # calculate ci width as a proportion of observed range
  mutate(ci_width_proportion = ci_width / range) %>%
  mutate(DV_type = "PI scores")

# combine
data_ci_width_proportions_combined <- 
  bind_rows(data_ci_width_proportions_D,
            data_ci_width_proportions_PI) %>%
  mutate(domain = as.factor(domain),
         method = fct_relevel(method, "bca", "basic", "normal", "percent"),
         trial_type = fct_relevel(trial_type, "tt1", "tt2", "tt3", "tt4"))

```

## Compare CI bootstrapping methods

```{r}

data_ci_width_proportions_combined_method <- 
  data_ci_width_proportions_D %>%
  mutate(domain = as.factor(domain),
         method = as.factor(method),
         method = fct_relevel(method, "bca", "basic", "normal", "percent"),
         trial_type = fct_relevel(trial_type, "tt1", "tt2", "tt3", "tt4")) %>%
  select(unique_id, domain, trial_type, method, ci_width_proportion) 

# fit model
fit_ci_width_proportions_method <- 
  lmer(ci_width_proportion ~ 1 + method + (1 | domain/trial_type), 
       data = data_ci_width_proportions_combined_method)

# results table
tab_model(fit_ci_width_proportions_method, 
          ci.hyphen = ", ")

# # summary
# summary(fit_ci_width_proportions_method)
#
# # plot re
# plot_model(fit_ci_width_proportions_method, 
#            colors = "bw",
#            type = "re") +
#   mdthemes::md_theme_linedraw() +
#   ggtitle("")

# extract re Tau
results_re_tau_ci_width_proportions_method <- fit_ci_width_proportions_method %>%
  merTools::REsdExtract() %>%
  as_tibble(rownames = "trial_type") %>%
  rename(tau = value) 

# extract marginal means
results_emm_ci_width_proportions_method <- 
  summary(emmeans(fit_ci_width_proportions_method, ~ method)) %>%
  dplyr::select(method, estimate = emmean, se = SE, ci_lower = asymp.LCL, ci_upper = asymp.UCL)

# combine
results_ci_width_proportions_method <- results_emm_ci_width_proportions_method %>%
  mutate(pi_lower = estimate - (1.96 * sqrt(se^2 + sum(results_re_tau_ci_width_proportions_method$tau)^2)),  # as in metafor package's implementation of credibility intervals, see metafor::predict.rma.R 
         pi_upper = estimate + (1.96 * sqrt(se^2 + sum(results_re_tau_ci_width_proportions_method$tau)^2))) 

# plot
p_ci_width_proportion_observed_range_method <- 
  results_ci_width_proportions_method %>%
  mutate(method = fct_rev(method)) %>%
  ggplot(aes(method, estimate, 
             color = method, 
             shape = method, 
             group = method)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pi_upper), position = position_dodge(width = 0.8), linetype = "dotted") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size = 1, position = position_dodge(width = 0.8)) +
  geom_point(position = position_dodge(width = 0.8), size = 2.5) +
  mdthemes::md_theme_linedraw() +
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), labels = c("0.00<br/>(Better)", "0.25", "0.50", "0.75", "1.00<br/>(Worse)")) +
  scale_shape_manual(labels = c("Percentile", "Normal", "Basic", "BCA"), values = c(17, 16, 15, 19)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7, labels = c("Percentile", "Normal", "Basic", "BCA"), direction = -1) +
  labs(shape = "Bootstrapping method",
       color = "Bootstrapping method",
       x = "",
       y = "Proportion of observed range covered <br/>by individual participants' 95% CIs") +
  theme(legend.position = "none") +
  coord_flip(ylim = c(0, 1))

p_ci_width_proportion_observed_range_method

results_ci_width_proportions_method %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

## Compare scoring method

```{r}

# fit model
fit_ci_width_proportions_scoring <- 
  data_ci_width_proportions_combined %>%
  filter(method == "basic" & 
           DV_type %in% c("*D* scores", "PI scores")) %>%
  lmer(ci_width_proportion ~ 1 + DV_type + (1 | domain/trial_type), 
       data = .)

# results table
tab_model(fit_ci_width_proportions_scoring,
          ci.hyphen = ", ")

# extract re Tau
results_re_tau_ci_width_proportions_scoring <-
  merTools::REsdExtract(fit_ci_width_proportions_scoring) %>%
  as_tibble(rownames = "trial_type") %>%
  rename(tau = value)

# extract marginal means
results_emm_ci_width_proportions_scoring <-
  summary(emmeans(fit_ci_width_proportions_scoring, ~ DV_type)) %>%
  dplyr::select(DV_type, estimate = emmean, se = SE, ci_lower = asymp.LCL, ci_upper = asymp.UCL)

# combine
results_ci_width_proportions_scoring <- results_emm_ci_width_proportions_scoring %>%
  mutate(pi_lower = estimate - (1.96 * sqrt(se^2 + sum(results_re_tau_ci_width_proportions_scoring$tau)^2)),  # as in metafor package's implementation of credibility intervals, see metafor::predict.rma.R
         pi_upper = estimate + (1.96 * sqrt(se^2 + sum(results_re_tau_ci_width_proportions_scoring$tau)^2)))

# plot
p_ci_width_proportion_observed_range_scoring <-
  ggplot(results_ci_width_proportions_scoring, aes(DV_type, estimate, color = DV_type, shape = DV_type, group = DV_type)) +
  geom_linerange(aes(ymin = pi_lower, ymax = pi_upper), position = position_dodge(width = 0.8), linetype = "dotted") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size = 1, position = position_dodge(width = 0.8)) +
  geom_point(position = position_dodge(width = 0.8), size = 2.5) +
  scale_shape_discrete(labels = c("*D* scores", "PI scores")) +
  scale_y_continuous(breaks = c(0, .25, .5, .75, 1), labels = c("0.00<br/>(Better)", "0.25", "0.50", "0.75", "1.00<br/>(Worse)")) +
  scale_shape_manual(labels = c("*D* scores", "PI scores"), values = c(15, 16)) +
  scale_color_viridis_d(begin = 0.3, end = 0.7, labels = c("*D* scores", "PI scores")) +
  mdthemes::md_theme_linedraw() +
  labs(x = "",
       y = "Proportion of observed range covered <br/>by individual participants' 95% CIs",
       color = "Scoring method",
       shape = "Scoring method") +
  theme(legend.position = "top") +
  coord_flip(ylim = c(0, 1))

p_ci_width_proportion_observed_range_scoring

results_ci_width_proportions_scoring %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Combined plots

## Plot 1

Plot 1 is merely illustrative. It shows the bootstrapped CIs for all participants, split by domain, but not splitting by trial type. *D* scores and the BCA method are displayed as they are used for the subsequent analyses. 

```{r fig.height=8, fig.width=8}

p_cis_by_domain

ggsave(filename  = "plots/figure_1_cis_by_domain.pdf",
       plot      = p_cis_by_domain,
       device    = "pdf",
       # path      = NULL,
       # dpi       = 300,
       units     = "in",
       width     = 8,
       height    = 8,
       limitsize = TRUE)

```

## Plot 2

Most probable CI width for *D* scores when bootstrapped using four different methods. Very similar results are found across methods. Overall Maximum A-Posteroi width (MAP, i.e., the mode of a continuous variable) was *D* = 1.31 +/- 0.01 between bootstrapping methods. Some domains and trial types did demonstrate smaller most probable widths. 

NB I elected not to meta-analyze these widths as they demonstrate very large skew at the individual level, which violate the assumptions of linear meta-analysis and underestimate the typical width (ie estimated mean widths << MAP observed widths). Rather than meta analyze, I simply report the domain and trial type level MAP values. More informative and valid analyses are presented below - ones which can directly compare the *D* and PI as an alternative. That could not be accomplished with a direct comparison with *D*/PI scores' 95% CIs as they are on different scales and follow different distributions. 

Given the minimal differences between bootstrapping methods, I employ the Bias Corrected and Accelerated methods (BCA) for all subsequent primary analyses. Sensitivity analyses that compare between bootstrapping methods are computed but not included in the main manuscript.


```{r fig.height=6, fig.width=6}

p_ci_widths

ggsave(filename  = "plots/figure_2_ci_widths.pdf",
       plot      = p_ci_widths,
       device    = "pdf",
       # path      = NULL,
       # dpi       = 300,
       units     = "in",
       width     = 8,
       height    = 6,
       limitsize = TRUE)

```

## Plot 3

The results of three hierarchical/meta analytic models are presented below, all of which provide information via different methods regarding how informative an individual participants' *D* (or PI) score is in terms of being able to state that they demonstrated evidence of a bias/IRAP effect/implicit attitude, whether that individual can be discriminated from other individuals in the same domain using the same trial type, and how much of the range of observed scores an individuals score's CI spans.  

(1) a meta-analysis of the proportion of participants that show non-zero scores (i.e., whose *D* or PI scores' 95% CIs exclude zero), (2) the proportion of scores that are discriminable from one another. Pairwise comparisons are made between all participants' scores within a domain and trial type, and the proportion that lie outside one another's CIs can be inferred to be different from one another (i.e., are discriminable as different from one another). This analysis is useful because it avoids the issue or debate around how meaningful a score's zero point is (i.e., *D* = 0, PI = 0.50) or whether it is a meaningful reference point. That is, previous research has argued that *D* = 0 cannot be inferred to represent no bias. Discriminability is agnostic to any individual comparison point and avoids this issue. (3) A meta analysis of the ratio between each participants' score and the maximium observed range of scores for that domain and trial type. I.e., given the observed range of scores across participants, what proportion of that range did each individual participant's score's Confidence Interval span. If each individual's CIs are compatible with a large proportion of the total range of all observed socres, then each score is so poorly estimated as to tell us very little about where on the spectrum each participant lies. 

All meta analyses compare *D* and PI scores to assess whether the results are dependent on the *D* algorithm which has been argued to be suboptimal. That is, I assess whether this issue can be trivially resolved by scoring the data a different way.

Note that the theoretical max possible range of D scores is -2 to +2, but such extreme scores are practically impossible. As such, in order to understand the CI width in terms of realistic data - and also in order to compare *D* and PI which are on different scales and distributions - I standardize CI widths by the observed range of scores for each domain and trial type. 

```{r fig.height=7, fig.width=5}

p_combined <- 
  p_prop_nonzero_scoring + 
  p_prop_discriminable_scoring + 
  p_ci_width_proportion_observed_range_scoring +
  plot_layout(ncol = 1, guides = "collect") & theme(legend.position = "top")

p_combined

ggsave(filename  = "plots/figure_3_metaanalyses.pdf",
       plot      = p_combined,
       device    = "pdf",
       # path      = NULL,
       # dpi       = 300,
       units     = "in",
       width     = 5,
       height    = 7,
       limitsize = TRUE)

```

# Session info

```{r}

sessionInfo()

```
