---
title: "Simulate trial-level IRAP data"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      cache.lazy=FALSE)
```

```{r}

# dependencies
library(tidyverse)
library(knitr)
library(kableExtra)
library(faux)

# set seed for reproducibility
set.seed(42)

# options
options(knitr.table.format = "html") # necessary configuration of tables

# disable scientific notation
options(scipen = 999)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, round, digits = n_digits)
}

```

# Parameterize IRAP data

```{r}

# get data from evaluative IRAPs
data_trial_level <- read_csv("../data/data_trial_level.csv") %>%
  filter(timepoint == "baseline") %>%
  group_by(domain, unique_id, block_type, trial_type) %>%
  mutate(trial_n = row_number()) %>%
  ungroup()

# outliers
data_outliers <- data_trial_level %>%
  distinct(unique_id, .keep_all = TRUE) %>%
  dplyr::select(unique_id, domain, mean_rt) %>%
  mutate(median_mean_rt = median(mean_rt, na.rm = TRUE),
         mad_mean_rt = mad(mean_rt, na.rm = TRUE)) %>%
  # exclude median +- 2MAD
  mutate(rt_outlier = ifelse(mean_rt < median_mean_rt - mad_mean_rt * 2 |
                               mean_rt > median_mean_rt + mad_mean_rt * 2, TRUE, FALSE)) %>%
  filter(rt_outlier == FALSE) %>%
  dplyr::select(unique_id, rt_outlier) %>%
  full_join(data_trial_level, by = "unique_id") %>%
  mutate(rt_outlier = ifelse(is.na(rt_outlier), TRUE, rt_outlier))

data_outliers_removed <- data_outliers %>%
  filter(rt_outlier == FALSE)

# # trim RTs>10000 ms, as part of D scoring
# data_trimmed <- data_outliers_removed %>%
#   dplyr::select(unique_id, domain, trial_type, trial_n, rt, block_type) %>%
#   filter(rt <= 10000 & rt > 0)

# trim RTs>10000 ms, as part of D scoring
data_trimmed <- data_outliers_removed %>%
  dplyr::select(unique_id, domain, trial_type, trial_n, rt, block_type) %>%
  filter(rt > 500 & rt < 10000) %>%
  mutate(log_rt = log(rt))

ggplot(data_trimmed, aes(log_rt, color = block_type)) +
  geom_density() +
  facet_wrap(~ trial_type)

```


```{r}

# # paramaterize real data
# data_param <- data_trimmed %>%
#   group_by(unique_id, domain, trial_type) %>%
#   summarize(mean_con = mean(rt[block_type == "con"], na.rm = TRUE),
#             mean_incon = mean(rt[block_type == "incon"], na.rm = TRUE),
#             sd_con = sd(rt[block_type == "con"], na.rm = TRUE),
#             sd_incon = sd(rt[block_type == "incon"], na.rm = TRUE)) %>%
#   ungroup() %>%
#   mutate(log_sd_con = log(sd_con),
#          log_sd_incon = log(sd_incon))

# paramaterize real data
data_param <- data_trimmed %>%
  group_by(unique_id, domain, trial_type) %>%
  summarize(mean_con = mean(log_rt[block_type == "con"], na.rm = TRUE),
            mean_incon = mean(log_rt[block_type == "incon"], na.rm = TRUE),
            sd_con = sd(log_rt[block_type == "con"], na.rm = TRUE),
            sd_incon = sd(log_rt[block_type == "incon"], na.rm = TRUE)) %>%
  ungroup() 
  # mutate(log_sd_con = log(sd_con),
  #        log_sd_incon = log(sd_incon))

# table
data_param %>%
  gather(metric, value, c(mean_con,
                          mean_incon,
                          sd_con,
                          sd_incon)) %>% 
                          #log_sd_con,
                          #log_sd_incon)) %>% 
  group_by(metric) %>%
  summarize(mean = mean(value),
            sd = sd(value)) %>%
  #round_df(1) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

# visualize
data_param %>%
  gather(block, mean, c("mean_con", "mean_incon")) %>%
  ggplot(aes(mean, color = block)) +
  geom_density()

# data_param %>%
#   gather(block, log_sd, c("log_sd_con", "log_sd_incon")) %>%
#   ggplot(aes(log_sd, color = block)) +
#   geom_density()

data_param %>%
  gather(block, sd, c("sd_con", "sd_incon")) %>%
  ggplot(aes(sd, color = block)) +
  geom_density()

ggplot(data_param, aes(mean_con, sd_con)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data_param, aes(mean_incon, sd_incon)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data_param, aes(mean_con, mean_incon)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data_param, aes(sd_con, sd_incon)) +
  geom_point() +
  geom_smooth(method = "lm")

```

# Level 1 - simulate multivariate normal distribution of means and standard deviations by block

```{r fig.height=5, fig.width=5}

n_simulated_participants <- 250

# sim level 1
simiulate_parameters_from_data <- function(data){
  data %>%
    dplyr::select(mean_con, 
                  mean_incon, 
                  sd_con,
                  sd_incon) %>%
    MASS::mvrnorm(n = n_simulated_participants,
                  mu = colMeans(.), 
                  Sigma = as.matrix(cov(.)), 
                  empirical = TRUE) %>%
    as.data.frame() %>%
    # mutate(sd_con = exp(log_sd_con),
    #        sd_incon = exp(log_sd_incon),
    #        id = row_number()) %>%
    mutate(id = row_number()) %>%
    dplyr::select(id, 
                  mean_con, 
                  mean_incon, 
                  sd_con, 
                  sd_incon)
                  #log_sd_con,
                  #log_sd_incon)
}

data_sim_level_1 <- data_param %>%
  group_by(domain, trial_type) %>%
  do(simiulate_parameters_from_data(.)) %>%
  mutate(id = paste(domain, id)) %>%
  arrange(id, domain, trial_type)

```

## Plot

```{r fig.height=5, fig.width=5}

ggplot(data_sim_level_1, aes(mean_con, sd_con)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm")

ggplot(data_sim_level_1, aes(mean_incon, sd_incon)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm")

ggplot(data_sim_level_1, aes(mean_con, mean_incon)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm")

ggplot(data_sim_level_1, aes(sd_con, sd_incon)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm")

```

- Both diff in means and log SD are normally distributed; therefore covariance matrix parameterizes them appropriately. 

# Level 2 - Use means and SDs to simulate trial level data that meets these parameters

trim any simulated participants with implausibly short simulated RTs. only .5% of RTs in the real data are less than 750 MS, so ditch these. can't trim individual negative values or the data wont match the parameters, so ditch them wholesale

```{r}

# # percent of RTs in real data less than 750 MS
# data_trimmed %>%
#   filter(rt > 750) %>%
#   count() %>%
#   pull(n) /
#   data_trimmed %>%
#   count() %>%
#   pull(n)

data_sim_level_1_tidied <- data_sim_level_1 %>%
  #dplyr::select(-log_sd_con, -log_sd_incon) %>%
  pivot_longer(names_to = "metric",
               values_to = "value",
               cols = c(mean_con, mean_incon, sd_con, sd_incon)) %>%
  separate(col = metric, into = c("metric", "block_type")) %>%
  pivot_wider(names_from = "metric",
              values_from = "value") 

# generate data from a normal distribution that conforms to a known mean and SD
# from https://stackoverflow.com/questions/18919091/generate-random-numbers-with-fixed-mean-and-sd
# results <- rnorm_empirical_basic(n = 18, mean = 1395, sd = 567)
# mean(results$simulated_value)
# sd(results$simulated_value)
rnorm_empirical <- function(n, mean, sd) { 
  mean + sd * scale(rnorm(n))
}


simiulate_data_from_simulated_parameters <- function(data, n_simulated_trials) {
  data %>%
    # take each row and simulate multiple new rows from it
    # solution https://blog.az.sg/posts/map-and-walk/
    pmap_dfr(function(...) {
      current <- tibble(...)
      tibble(id = current$id, 
             domain = current$domain,
             trial_type = current$trial_type,
             block_type = current$block_type,
             simulated_value = rnorm_empirical(mean = current$mean, 
                                               sd = current$sd, 
                                               n = n_simulated_trials)) 
    }) 
}

simulated_data_18_temp <- 
  simiulate_data_from_simulated_parameters(data = data_sim_level_1_tidied, n_simulated_trials = 18) 

simulated_data_180_temp <- 
  simiulate_data_from_simulated_parameters(data = data_sim_level_1_tidied, n_simulated_trials = 180) 

# exponentiate back to millisecond scale ie RTs
simulated_data_18 <- simulated_data_18_temp %>%
  mutate(simulated_value = exp(simulated_value))

simulated_data_180 <- simulated_data_180_temp %>%
  mutate(simulated_value = exp(simulated_value))

```

## Test

### 18 trials (as in real data)

```{r}

simulated_data_18 %>%
  distinct(id, .keep_all = TRUE) %>%
  count(domain) %>%
  count(n)

simulated_data_18_summary <- simulated_data_18 %>%
  group_by(id, block_type) %>%
  summarize(sim_mean = mean(simulated_value),
            sim_sd = sd(simulated_value),
            .groups = "drop") %>%
  ungroup() %>%
  left_join(data_sim_level_1_tidied, by = c("id", "block_type"))

ggplot(simulated_data_18_summary, aes(mean, sim_mean)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~ block_type)

ggplot(simulated_data_18_summary, aes(sd, sim_sd)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~ block_type)

ggplot(data_trimmed, aes(rt, color = block_type)) +
  geom_density(alpha = 0.1) +
  facet_wrap(~ trial_type) +
  xlim(0, 5000) +
  ggtitle("Real IRAP data")

ggplot(simulated_data_18, aes(simulated_value, color = block_type)) +
  geom_density(alpha = 0.1) +
  facet_wrap(~ trial_type) +
  xlim(0, 5000) +
  ggtitle("Simulated IRAP data (18 trials per D score)")

```

### 180 trials

```{r}

simulated_data_180 %>%
  distinct(id, .keep_all = TRUE) %>%
  count(domain) %>%
  count(n)

simulated_data_180_summary <- simulated_data_180 %>%
  group_by(id, block_type) %>%
  summarize(sim_mean = mean(simulated_value),
            sim_sd = sd(simulated_value),
            .groups = "drop") %>%
  ungroup() %>%
  left_join(data_sim_level_1_tidied, by = c("id", "block_type"))

ggplot(simulated_data_180_summary, aes(mean, sim_mean)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~ block_type)

ggplot(simulated_data_180_summary, aes(sd, sim_sd)) +
  geom_point(alpha = 0.1) +
  facet_wrap(~ block_type)

ggplot(data_trimmed, aes(rt, color = block_type)) +
  geom_density(alpha = 0.1) +
  facet_wrap(~ trial_type) +
  xlim(0, 5000) +
  ggtitle("Real IRAP data")

ggplot(simulated_data_180, aes(simulated_value, color = block_type)) +
  geom_density(alpha = 0.1) +
  facet_wrap(~ trial_type) +
  xlim(0, 5000) +
  ggtitle("Simulated IRAP data (180 trials per D score)")

```

# Write to disk

```{r}

# write_csv(simulated_data_18,  "../data/data_trial_level_simulated_18_trials.csv")
# write_csv(simulated_data_180, "../data/data_trial_level_simulated_180_trials.csv")

write_rds(simulated_data_18,  "../data/data_trial_level_simulated_18_trials.rds", compress = "gz")
write_rds(simulated_data_180, "../data/data_trial_level_simulated_180_trials.rds", compress = "gz")

```





